---
title: "View Thornthwaite Budgets as Maps"
output: html_notebook
---

# Load Libraries
```{r}

# load libraries

library(package = "tidyverse")
library(package = "lubridate")
library(package = "tidync")
library(package = "ncdf4")
library(package = "ncmeta")
#library(package = "RCurl")

library(package = "usmap")


```


# Get OS

```{r}
# Get OS

OS = Sys.info()[1]

print(OS)

if (OS == "Darwin")
{
  root_path = "~/GitHub/LOCA_CHEYENNE_THORNTHWAITE_MAP_WEB/" 
} else {
  root_path = "~/GitHub/LOCA_CHEYENNE_THORNTHWAITE_MAP_WEB/" 
}

```



# Get User Controlled Parameters


## List of Climate Periods

| 10-y Climate Period Duration | 30-y Climate Period Duration |
|------------------------------|------------------------------|
|        1950-1959             |          1950-1979           |
|        1960-1969             |          1970-1999           |
|        1970-1979             |          1976-2005           |
|        1980-1989             |          2000-2029           |
|        1990-1999             |          2011-2040           |
|        2000-2009             |          2036-2065           |
|        2010-2019             |          2061-2090           |
|        2020-2029             |          2070-2099           |
|        2030-2039             |                              |
|        2040-2049             |                              |
|        2050-2059             |                              |
|        2060-2069             |                              |
|        2070-2079             |                              |
|        2080-2089             |                              |
|        2090-2099             |                              |

## List of Variables

```{r}

# Get Variable List and LUT

variable_metadata_file = str_c(root_path,
                               "Thornthwaite_Budget_Parameters.RData",
                               sep = "")

load(file = variable_metadata_file)


print(thornthwaite_variable_list)


```


##  User Controlled Parameters Entered
       
```{r}

# Get User Controlled Parameters - Emulating Pull-down Menus

period_length = "30" # climatology period years

selected_variable = "monthly_precipitation"



baseline_period = "1976-2005"

examined_period = "2036-2065"


```

# set Netcdf Import Variables

```{r}

# set Netcdf Import Variables

target_variable = str_c("mean_",
                        selected_variable,
                        sep = "")

stdev_variable =  str_c("stdev_",
                        selected_variable,
                        sep = "")


```


# set color tables

```{r}




# colour palate selection

# default 


ensavg_palette  = "Spectral"
ensavg_direction = 1

anom_palette    = "BrBG"
anom_direction   = 1






if (str_detect(string = selected_variable, pattern = "emperature")) {
  ensavg_palette  = "Spectral"
  ensavg_direction = 1

  anom_palette    = "RdBu"
  anom_direction   = -1
}


# drying (deficit)

if (str_detect(string = selected_variable, pattern = "deficit") | 
    str_detect(string = selected_variable, pattern = "evap") ) {
  ensavg_palette   = "YlOrRd"
  ensavg_direction = 1
  
  anom_palette     = "BrBG"
  anom_direction   = -1  
}


# recharge & surplus & storage 

if (str_detect(string = selected_variable, pattern = "recharge") |
    str_detect(string = selected_variable, pattern = "surplus")  |
    str_detect(string = selected_variable, pattern = "storage")  |
    str_detect(string = selected_variable, pattern = "recipit")) {

  ensavg_palette   = "YlGnBu"
  ensavg_direction = 1
  
  anom_palette     = "BrBG"
  anom_direction   = 1 
}

# snowmelt 

if (str_detect(string = selected_variable, pattern = "snow"))  {
  ensavg_palette   = "PuBu"
  ensavg_direction = 1
  
  anom_palette     = "RdBu"
  anom_direction   = 1 
}



```

## Pull Available Period Dates

```{r}

# Get the calendar values

period_file_location = str_c(root_path,
                             "Cheyenne_climatology_years_",
                             period_length,
                             "-yr.RData",
                             sep="") 

load(file = period_file_location)



print(climatology_period_info)

```



# Make a DEM Reference Map

```{r}

# Make a DEM Reference Map

dem_file_location = str_c(root_path,
                          "gtopo30_016_Cheyenne.RData",
                          sep="") 

load(file = dem_file_location)

# Our domain boundary for our maps

min_lat =   min(dem_frame$lat) 
max_lat =   max(dem_frame$lat) 

min_lon  =   min(dem_frame$lon) 
max_lon  =   max(dem_frame$lon) 




# Cracking the Natural Map Boundary Set for the first-level subdivisions of 
# the US and Canada (states and provinces)
  
us_states = map_data(map = "state")
counties  = map_data(map = "county")

# Plot
  
ggplot(data = dem_frame)+
  aes(x    = lon, 
      y    = lat, 
      fill = elevation) +
  theme_bw() +
  labs(title    = "Cheyenne Basin",
       subtitle =  dem_long_name,
       fill      = "m") +
  xlab(label = "Longitude") +
  ylab(label = "Latitude") +
  coord_fixed(xlim    = c(min_lon, max_lon),
              ylim    = c(min_lat, max_lat) ,
              ratio   = 1/cos(pi/180 * mean(dem_frame$lat)) ) +
  scale_fill_gradientn(colors     = terrain.colors(10),
                       space      = "Lab",
                       na.value   = "white",
                       aesthetics = "fill",
                       limits     = c(min(dem_frame$elevation),
                                      max(dem_frame$elevation))) +
  geom_raster()  +
  geom_polygon(data    = us_states,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               fill    = NA, 
               size    = 1.0, 
              color   = "black")  +
  geom_polygon(data    = counties,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               size    = 0.25, 
               fill    = NA, 
               color   = "black")  

Lon = unique(dem_frame$lon)
Lat = unique(dem_frame$lat)


```


# Make an Available Water Capacity Reference Map

```{r}

# Make a DEM Reference Map

awc_file_location =  str_c(root_path,
                          "AWC_Cheyenne.RData",
                          sep="") 

load(file = awc_file_location)

# Our domain boundary for our maps

awc_lat =   min(awc_frame$lat) 
awc_lat =   max(awc_frame$lat) 

awc_lon  =   min(awc_frame$lon) 
awc_lon  =   max(awc_frame$lon) 




# Cracking the Natural Map Boundary Set for the first-level subdivisions of 
# the US and Canada (states and provinces)
  
us_states = map_data(map = "state")
counties  = map_data(map = "county")

# Plot
  
ggplot(data = awc_frame)+
  aes(x    = lon, 
      y    = lat, 
      fill = usda_awc) +
  theme_bw() +
  labs(title    = "Cheyenne Basin",
       subtitle =  "Max Available Soil Water Storage",
       fill      = "mm") +
  xlab(label = "Longitude") +
  ylab(label = "Latitude") +
  coord_fixed(xlim    = c(min_lon, max_lon),
              ylim    = c(min_lat, max_lat) ,
              ratio   = 1/cos(pi/180 * mean(awc_frame$lat)) ) +
  scale_fill_distiller(palette    = "YlGnBu",
                       space      = "Lab",
                       direction  = 1,
                       na.value   = "white",
                       guide      = "colourbar",
                       aesthetics = "fill",
                       limits     = c(0,max(awc_frame$usda_awc))) +
  geom_raster()  +
  geom_polygon(data    = us_states,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               fill    = NA, 
               size    = 1.0, 
              color   = "black")  +
  geom_polygon(data    = counties,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               size    = 0.25, 
               fill    = NA, 
               color   = "black")  



```
# Crack the Big NetCDF File for one parameter

```{r}

# Crack the Big NetCDF File for one parameter
     
data_file_location = str_c(root_path,
                          "/local_climatologies/",
                          "LOCA_THORNTHWAITE_ALL_ENSEMBLES_CLIM_",
                           period_length,
                           "-yr.nc", 
                           sep = "")


variable_frame = tidync(x = data_file_location)  %>% 
  activate(target_variable, 
           stdev_variable) %>% 
  hyper_tibble(select_var = c(target_variable,
                              stdev_variable)) %>%
  rename(value = target_variable,
         stdv  = stdev_variable) %>%
  filter(!is.na(value)) %>%
  mutate(scenario           = as.factor(scenario),
         ensemble           = as.factor(ensemble),
         climatology_period = as.factor(climatology_period),
         month              = fct_recode(as.factor(month), 
                                         January   =  "1",
                                         February  =  "2",
                                         March     =  "3",
                                         April     =  "4",
                                         May       =  "5",
                                         June      =  "6",
                                         July      =  "7",
                                         August    =  "8",
                                         September =  "9",
                                         October   = "10",
                                         November  = "11",
                                         December  = "12") )


Ensembles = levels(variable_frame$ensemble)
Scenarios = levels(variable_frame$scenario)
Period    = levels(variable_frame$climatology_period)
Month     = levels(variable_frame$month)

variable_metadata = thornthwaite_variable_list %>%
  filter(variable == selected_variable)



```

# Setup Data for Comparisons

```{r}



# Setup Data for Comparisons


baseline_and_future_variable_frame = variable_frame %>% 
  filter((climatology_period == baseline_period) | 
         (climatology_period == examined_period) )  %>%
  mutate(scenario = ifelse(test = (scenario == "HIST_RCP45") & (climatology_period == baseline_period),
                           yes  = "HIST",
                           no   = ifelse(test = (scenario == "HIST_RCP45"),
                                         yes  = "RCP45",
                                         no   = "RCP85"))) %>%
  select(-c("climatology_period","stdv")) %>%
  mutate(baseline_period = baseline_period,
         examined_period = examined_period)

baseline_and_future_variable_frame = spread(data  = baseline_and_future_variable_frame,
                                            key   = scenario,
                                            value = value)


baseline_and_future_variable_frame$anom_RCP45_m_HIST = baseline_and_future_variable_frame$RCP45 - baseline_and_future_variable_frame$HIST
baseline_and_future_variable_frame$anom_RCP85_m_HIST = baseline_and_future_variable_frame$RCP85 - baseline_and_future_variable_frame$HIST

```

# plot baseline internal variability

```{r}
# plot baseline internal variability

baseline_variance_frame = variable_frame %>% 
  filter(climatology_period == baseline_period)  %>%
  mutate(scenario = "HIST") %>%
  select(-c("climatology_period",
            "value",
            "scenario")) %>%
  group_by(lon,
           lat,
           month) %>%
  summarize(intern_var = mean(stdv)) %>%
  ungroup()



# plot baseline internal variability

ggplot(data = baseline_variance_frame)+
  aes(x    = lon, 
      y    = lat, 
      fill = intern_var) +
  theme_bw() +
  theme(strip.background = element_rect(fill=NA),
        axis.title.x     = element_blank(),
        axis.text.x      = element_blank(),
        axis.ticks.x     = element_blank(),
        axis.title.y     = element_blank(),
        axis.text.y      = element_blank(),
        axis.ticks.y     = element_blank()) +
  facet_wrap(facets = "month") +
  labs(title    = "Cheyenne Basin Thornthwaite Budgets",
       subtitle =  str_c(variable_metadata$long_name,
                         " Internal Variability;  Historical (",
                         baseline_period, 
                         ")",
                         sep = ""),
       caption = "CMIP5 LOCA Ensemble Means of σ",
       fill    = variable_metadata$units) +
  xlab(label = "Longitude") +
  ylab(label = "Latitude") +
  coord_fixed(xlim    = c(min_lon, max_lon),
              ylim    = c(min_lat, max_lat) ,
              ratio   = 1/cos(pi/180 * mean(dem_frame$lat)) ) +
  scale_fill_distiller(palette    = 'Spectral',
                       space      = "Lab",
                       direction  = -1,
                       na.value   = "white",
                       guide      = "colourbar",
                       aesthetics = "fill") +
  geom_raster()  +
  geom_polygon(data    = us_states,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               fill    = NA, 
               size    = 1.0, 
              color   = "black")  +
  geom_polygon(data    = counties,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               size    = 0.25, 
               fill    = NA, 
               color   = "black")  

```


# Setup Data for Comparisons

```{r}

# Setup Data for Comparisons

ensemble_average_data_frame = left_join(x  = baseline_and_future_variable_frame,
                                        y  = baseline_variance_frame,
                                        by = c("lon",
                                               "lat",
                                               "month"))
ensemble_average_data_frame = ensemble_average_data_frame %>%
  group_by(lon,
           lat,
           month,
           intern_var) %>%
  summarize(ensemble_mean_HIST  = mean(HIST),
            ensemble_mean_RCP45 = mean(RCP45),
            ensemble_mean_RCP85 = mean(RCP85),
            ensemble_mean_A4mH  = mean(anom_RCP45_m_HIST),
            ensemble_mean_A8mH  = mean(anom_RCP85_m_HIST),
            sign_A4mH           = sum( sign(mean(anom_RCP45_m_HIST)) == sign(anom_RCP45_m_HIST) )/n(),
            sign_A8mH           = sum( sign(mean(anom_RCP85_m_HIST)) == sign(anom_RCP85_m_HIST) )/n()) %>%
  ungroup()   

ensemble_average_data_frame = ensemble_average_data_frame %>%
  mutate(baseline_period = baseline_period,
         examined_period = examined_period,
         frac_var_A4mH   = abs(ensemble_mean_A4mH)/intern_var,
         frac_var_A8mH   = abs(ensemble_mean_A8mH)/intern_var)



ensemble_mean_value_max = ceiling(max(c(ensemble_average_data_frame$ensemble_mean_HIST,
                                        ensemble_average_data_frame$ensemble_mean_RCP45,
                                        ensemble_average_data_frame$ensemble_mean_RCP85)))

ensemble_mean_value_min = floor(min(c(ensemble_average_data_frame$ensemble_mean_HIST,
                                       ensemble_average_data_frame$ensemble_mean_RCP45,
                                       ensemble_average_data_frame$ensemble_mean_RCP85)))
                                
ensemble_mean_value_limits = c(ensemble_mean_value_min,ensemble_mean_value_max)



ensemble_mean_anom_abs_max  = ceiling(max(c(abs(ensemble_average_data_frame$ensemble_mean_A4mH),
                                            abs(ensemble_average_data_frame$ensemble_mean_A8mH))))


ensemble_mean_anom_limits = ensemble_mean_anom_abs_max * c(-1,1)



```

# 

## Plot Historical Period

```{r}



ggplot(data = ensemble_average_data_frame)+
  aes(x    = lon, 
      y    = lat, 
      fill = ensemble_mean_HIST) +
  theme_bw() +
  theme(strip.background = element_rect(fill=NA),
        axis.title.x     = element_blank(),
        axis.text.x      = element_blank(),
        axis.ticks.x     = element_blank(),
        axis.title.y     = element_blank(),
        axis.text.y      = element_blank(),
        axis.ticks.y     = element_blank()) +
  facet_wrap(facets = "month") +
  labs(title    = "Cheyenne Basin Thornthwaite Budgets",
       subtitle =  str_c(variable_metadata$long_name,
                         ";  Historical (",
                         baseline_period, 
                         ")",
                         sep = ""),
       caption = "CMIP5 LOCA Ensemble Means",
       fill    = variable_metadata$units) +
  xlab(label = "Longitude") +
  ylab(label = "Latitude") +
  coord_fixed(xlim    = c(min_lon, max_lon),
              ylim    = c(min_lat, max_lat) ,
              ratio   = 1/cos(pi/180 * mean(dem_frame$lat)) ) +
  scale_fill_distiller(palette    = ensavg_palette,
                       space      = "Lab",
                       direction  = ensavg_direction,
                       na.value   = "white",
                       guide      = "colourbar",
                       aesthetics = "fill",
                       limits     = ensemble_mean_value_limits) +
  geom_raster()  +
  geom_polygon(data    = us_states,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               fill    = NA, 
               size    = 0.5, 
              color   = "black")  +
  geom_polygon(data    = counties,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               size    = 0.2, 
               fill    = NA, 
               color   = "black")  


```
## RCP 4.5 

```{r}


ggplot(data = ensemble_average_data_frame)+
  aes(x    = lon, 
      y    = lat, 
      fill = ensemble_mean_RCP45) +
  theme_bw() +
  theme(strip.background = element_rect(fill=NA),
        axis.title.x     = element_blank(),
        axis.text.x      = element_blank(),
        axis.ticks.x     = element_blank(),
        axis.title.y     = element_blank(),
        axis.text.y      = element_blank(),
        axis.ticks.y     = element_blank()) +
  facet_wrap(facets = "month") +
  labs(title    = "Cheyenne Basin Thornthwaite Budgets",
       subtitle =  str_c(variable_metadata$long_name,
                         ";  RCP 4.5 (",
                         examined_period, 
                         ")",
                         sep = ""),
       caption = "CMIP5 LOCA Ensemble Means",
       fill    = variable_metadata$units) +
  xlab(label = "Longitude") +
  ylab(label = "Latitude") +
  coord_fixed(xlim    = c(min_lon, max_lon),
              ylim    = c(min_lat, max_lat) ,
              ratio   = 1/cos(pi/180 * mean(dem_frame$lat)) ) +
  scale_fill_distiller(palette    = ensavg_palette,
                       space      = "Lab",
                       direction  = ensavg_direction,
                       na.value   = "white",
                       guide      = "colourbar",
                       aesthetics = "fill",
                       limits     = ensemble_mean_value_limits) +
  geom_raster()  +
  geom_polygon(data    = us_states,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               fill    = NA, 
               size    = 0.5, 
              color   = "black")  +
  geom_polygon(data    = counties,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               size    = 0.2, 
               fill    = NA, 
               color   = "black")  



```

## RCP 8.5 

```{r}

ggplot(data = ensemble_average_data_frame)+
  aes(x    = lon, 
      y    = lat, 
      fill = ensemble_mean_RCP85) +
  theme_bw() +
  theme(strip.background = element_rect(fill=NA),
        axis.title.x     = element_blank(),
        axis.text.x      = element_blank(),
        axis.ticks.x     = element_blank(),
        axis.title.y     = element_blank(),
        axis.text.y      = element_blank(),
        axis.ticks.y     = element_blank()) +
  facet_wrap(facets = "month") +
  labs(title    = "Cheyenne Basin Thornthwaite Budgets",
       subtitle =  str_c(variable_metadata$long_name,
                         ";  RCP 8.5 (",
                         examined_period, 
                         ")",
                         sep = ""),
       caption = "CMIP5 LOCA Ensemble Means",
       fill    = variable_metadata$units) +
  xlab(label = "Longitude") +
  ylab(label = "Latitude") +
  coord_fixed(xlim    = c(min_lon, max_lon),
              ylim    = c(min_lat, max_lat) ,
              ratio   = 1/cos(pi/180 * mean(dem_frame$lat)) ) +
  scale_fill_distiller(palette    = ensavg_palette,
                       space      = "Lab",
                       direction  = ensavg_direction,
                       na.value   = "white",
                       guide      = "colourbar",
                       aesthetics = "fill",
                       limits     = ensemble_mean_value_limits) +
  geom_raster()  +
  geom_polygon(data    = us_states,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               fill    = NA, 
               size    = 0.5, 
              color   = "black")  +
  geom_polygon(data    = counties,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               size    = 0.2, 
               fill    = NA, 
               color   = "black")  



```

#  Plot Anomalies

# RCP 4.5


```{r}

# RCP 4.5


ggplot(data = ensemble_average_data_frame)+
  aes(x    = lon, 
      y    = lat, 
      fill = ensemble_mean_A4mH) +
  theme_bw() +
  theme(strip.background = element_rect(fill=NA),
        axis.title.x     = element_blank(),
        axis.text.x      = element_blank(),
        axis.ticks.x     = element_blank(),
        axis.title.y     = element_blank(),
        axis.text.y      = element_blank(),
        axis.ticks.y     = element_blank()) +
  facet_wrap(facets = "month") +
  labs(title    = "Cheyenne Basin Thornthwaite Budget Anomalies",
       subtitle =  str_c(variable_metadata$long_name,
                         ";  RCP 4.5 (",
                         examined_period, 
                         ") - (",
                         baseline_period,
                         ")",
                         sep = ""),
       caption = "CMIP5 LOCA Ensemble Means",
       fill    = variable_metadata$units) +
  xlab(label = "Longitude") +
  ylab(label = "Latitude") +
  coord_fixed(xlim    = c(min_lon, max_lon),
              ylim    = c(min_lat, max_lat) ,
              ratio   = 1/cos(pi/180 * mean(dem_frame$lat)) ) +
  scale_fill_distiller(palette    = anom_palette,
                       space      = "Lab",
                       direction  = anom_direction,
                       na.value   = "white",
                       guide      = "colourbar",
                       aesthetics = "fill",
                       limits     = ensemble_mean_anom_limits) +
  geom_raster()  +
  geom_polygon(data    = us_states,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               fill    = NA, 
               size    = 0.5, 
              color   = "black")  +
  geom_polygon(data    = counties,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               size    = 0.2, 
               fill    = NA, 
               color   = "black")  



```


# RCP 8.5


```{r}

# RCP 8.5


ggplot(data = ensemble_average_data_frame)+
  aes(x    = lon, 
      y    = lat, 
      fill = ensemble_mean_A8mH) +
  theme_bw() +
  theme(strip.background = element_rect(fill=NA),
        axis.title.x     = element_blank(),
        axis.text.x      = element_blank(),
        axis.ticks.x     = element_blank(),
        axis.title.y     = element_blank(),
        axis.text.y      = element_blank(),
        axis.ticks.y     = element_blank()) +
  facet_wrap(facets = "month") +
  labs(title    = "Cheyenne Basin Thornthwaite Budget Anomalies",
       subtitle =  str_c(variable_metadata$long_name,
                         ";  RCP 8.5 (",
                         examined_period, 
                         ") - (",
                         baseline_period,
                         ")",
                         sep = ""),
       caption = "CMIP5 LOCA Ensemble Means",
       fill    = variable_metadata$units) +
  xlab(label = "Longitude") +
  ylab(label = "Latitude") +
  coord_fixed(xlim    = c(min_lon, max_lon),
              ylim    = c(min_lat, max_lat) ,
              ratio   = 1/cos(pi/180 * mean(dem_frame$lat)) ) +
  scale_fill_distiller(palette    = anom_palette,
                       space      = "Lab",
                       direction  = anom_direction,
                       na.value   = "white",
                       guide      = "colourbar",
                       aesthetics = "fill",
                       limits     = ensemble_mean_anom_limits) +
  geom_raster()  +
  geom_polygon(data    = us_states,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               fill    = NA, 
               size    = 0.5, 
              color   = "black")  +
  geom_polygon(data    = counties,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               size    = 0.2, 
               fill    = NA, 
               color   = "black")  



```


