---
title: "View Thornthwaite Budgets as Mathergrams"
output: html_notebook
---

# Load Libraries
```{r}

# load libraries

library(package = "tidyverse")
library(package = "lubridate")
library(package = "tidync")
library(package = "ncdf4")
library(package = "ncmeta")
library(package = "RCurl")

library(package = "usmap")


```
# Get User Controlled Parameters


## List of Climate Periods

| 10-y Climate Period Duration | 30-y Climate Period Duration |
|------------------------------|------------------------------|
|        1950-1959             |          1950-1979           |
|        1960-1969             |          1970-1999           |
|        1970-1979             |          1976-2005           |
|        1980-1989             |          2000-2029           |
|        1990-1999             |          2011-2040           |
|        2000-2009             |          2036-2065           |
|        2010-2019             |          2061-2090           |
|        2020-2029             |          2070-2099           |
|        2030-2039             |                              |
|        2040-2049             |                              |
|        2050-2059             |                              |
|        2060-2069             |                              |
|        2070-2079             |                              |
|        2080-2089             |                              |
|        2090-2099             |                              |

## List of Variables

```{r}

# Get Variable List and LUT

variable_metadata_file = "~/GitHub/LOCA_CHEYENNE_THORNTHWAITE_WEB/Thornthwaite_Budget_Parameters.RData"

load(file = variable_metadata_file)


print(thornthwaite_variable_list)


```


##  User Controlled Parameters Entered
       
```{r}

# Get User Controlled Parameters - Emulating Pull-down Menus

period_length = "30" # climatology period years

# Latitude Longitude

target_latitude  =   44.09375
target_longitude = -103.21875

# Periods

baseline_period = "1976-2005"

examined_period = "2036-2065"

```

# set Netcdf Import Variables

```{r}

# set Netcdf Import Variables

variable = thornthwaite_variable_list$variable

target_variables = str_c("mean_",
                        thornthwaite_variable_list$variable,
                        sep = "")

stdev_variables =  str_c("stdev_",
                        thornthwaite_variable_list$variable,
                        sep = "")


```



## Pull Available Period Dates

```{r}

# Get the calendar values

period_file_location = str_c("~/GitHub/LOCA_CHEYENNE_THORNTHWAITE_WEB/Cheyenne_climatology_years_",
                             period_length,
                             "-yr.RData",
                             sep="") 

load(file = period_file_location)



print(climatology_period_info)

```



# Make a DEM Reference Map

```{r}

# Make a DEM Reference Map

dem_file_location = "~/GitHub/LOCA_CHEYENNE_THORNTHWAITE_WEB/gtopo30_016_Cheyenne.RData"

load(file = dem_file_location)

# Our domain boundary for our maps

min_lat =   min(dem_frame$lat) 
max_lat =   max(dem_frame$lat) 

min_lon  =   min(dem_frame$lon) 
max_lon  =   max(dem_frame$lon) 


Lon = unique(dem_frame$lon)
Lat = unique(dem_frame$lat)


# Cracking the Natural Map Boundary Set for the first-level subdivisions of 
# the US and Canada (states and provinces)
  
us_states = map_data(map = "state")
counties  = map_data(map = "county")

```
# Plot DEM
```{r}

# Plot
  
ggplot(data = dem_frame)+
  aes(x    = lon, 
      y    = lat, 
      fill = elevation) +
  theme_bw() +
  labs(title    = "Cheyenne Basin",
       subtitle =  dem_long_name,
       fill      = "m",
       caption = str_c("Coordinates: ",
                       target_longitude," °E, ",
                       target_latitude, " °N",
                       sep = "")) +
  xlab(label = "Longitude") +
  ylab(label = "Latitude") +
  coord_fixed(xlim    = c(min_lon, max_lon),
              ylim    = c(min_lat, max_lat) ,
              ratio   = 1/cos(pi/180 * mean(dem_frame$lat)) ) +
  scale_fill_gradientn(colors     = terrain.colors(10),
                       space      = "Lab",
                       na.value   = "white",
                       aesthetics = "fill",
                       limits     = c(min(dem_frame$elevation),
                                      max(dem_frame$elevation))) +
  geom_raster()  +
  geom_polygon(data    = us_states,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               fill    = NA, 
               size    = 1.0, 
              color   = "black")  +
  geom_polygon(data    = counties,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               size    = 0.25, 
               fill    = NA, 
               color   = "black")  +
  geom_vline(xintercept = target_longitude,
             color      = "white") +
  geom_hline(yintercept = target_latitude,
             color      = "white") 




```


# Make an Available Water Capacity Reference Map

```{r}

# Make a DEM Reference Map

awc_file_location = "~/GitHub/LOCA_CHEYENNE_THORNTHWAITE_WEB/AWC_Cheyenne.RData"

load(file = awc_file_location)



```

# Plot AWC

```{r}
# Plot AWC
  
ggplot(data = awc_frame)+
  aes(x    = lon, 
      y    = lat, 
      fill = usda_awc) +
  theme_bw() +
  labs(title    = "Cheyenne Basin",
       subtitle =  "Max Available Soil Water Storage",
       fill      = "mm",
       caption = str_c("Coordinates: ",
                       target_longitude," °E, ",
                       target_latitude, " °N",
                       sep = "")) +
  xlab(label = "Longitude") +
  ylab(label = "Latitude") +
  coord_fixed(xlim    = c(min_lon, max_lon),
              ylim    = c(min_lat, max_lat) ,
              ratio   = 1/cos(pi/180 * mean(awc_frame$lat)) ) +
  scale_fill_distiller(palette    = "YlGnBu",
                       space      = "Lab",
                       direction  = 1,
                       na.value   = "white",
                       guide      = "colourbar",
                       aesthetics = "fill",
                       limits     = c(0,max(awc_frame$usda_awc))) +
  geom_raster()  +
  geom_polygon(data    = us_states,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               fill    = NA, 
               size    = 1.0, 
              color   = "black")  +
  geom_polygon(data    = counties,
               mapping = aes(x     =   long, 
                             y     =    lat,
                             group =  group),
               size    = 0.25, 
               fill    = NA, 
               color   = "black")  +
  geom_vline(xintercept = target_longitude,
             color      = "white") +
  geom_hline(yintercept = target_latitude,
             color      = "white") 




```
# Crack the Big NetCDF File for one parameter

```{r}

# Crack the Big NetCDF File for one parameter
     
data_file_location = str_c("~/GitHub/LOCA_CHEYENNE_THORNTHWAITE_WEB/local_climatologies/LOCA_THORNTHWAITE_ALL_ENSEMBLES_CLIM_",
                           period_length,
                           "-yr.nc", 
                           sep = "")


variable_frame = tidync(x = data_file_location)  %>% 
  activate(target_variables) %>%
  hyper_tibble(select_var = target_variables,
               lat        = (lat == target_latitude),
               lon        = (lon == target_longitude)) %>%
  mutate(ensemble           = as.factor(ensemble),
         climatology_period = as.factor(climatology_period),
         month              = fct_recode(as.factor(month), 
                                         Jan  =  "1",
                                         Feb  =  "2",
                                         Mar  =  "3",
                                         Apr  =  "4",
                                         May  =  "5",
                                         Jne  =  "6",
                                         Jly  =  "7",
                                         Aug  =  "8",
                                         Sep  =  "9",
                                         Oct  = "10",
                                         Nov  = "11",
                                         Dec  = "12") ) 

variable_frame = left_join(x  = variable_frame,
                           y  = climatology_period_info,
                           by = "climatology_period")  %>%
  mutate(scenario = ifelse(test = scenario == "HIST_RCP45",
                           yes  = string_rcp45,
                           no   = "RCP 8.5")) %>%
  select(-c("string_rcp45"))

variable_frame$scenario = as.factor(variable_frame$scenario)

variable_frame = variable_frame %>%
  mutate(scenario_label = str_c(scenario,
                                " (",
                                climatology_period,
                                ")",
                                sep = ""))

variable_frame$scenario_label = as.factor(variable_frame$scenario_label)


variable_frame = gather(data   = variable_frame,
                        key   = "variable",
                        value = "value",
                        target_variables)

variable_frame = variable_frame %>%
  mutate(variable = str_sub(string = variable, 
                            start  = 6))



variable_frame$variable = fct_recode(variable_frame$variable, 
                                         `Deficit (mm)`  =  "deficit",
                                         `Evapotranspiration (mm)`  =  "evapotranspiration",
                                         `Monthly Precipitation (mm)`  =  "monthly_precipitation",
                                         `Monthly Temperature (°C)`  =  "monthly_temperature",
                                         `Potential Evaporation (mm)`  =  "potential_evapotranspiration",
                                         `Recharge (mm)`  =  "recharge",
                                         `Snowpack (mm)`  =  "snowpack",
                                         `Storage (mm)`  =  "storage",
                                         `Surplus (mm)`  =  "surplus") 


Ensembles = levels(variable_frame$ensemble)
Scenarios = levels(variable_frame$scenario)
Period    = levels(variable_frame$climatology_period)
Month     = levels(variable_frame$month)
Variable  = levels(variable_frame$variable)





```


# Reduce Data to Targeted Years


```{r}
# Reduce Data to Targeted Years

plotting_master_frame = variable_frame %>%
  filter((climatology_period == baseline_period) | 
         (climatology_period == examined_period)) %>%
  mutate(month_num = as.numeric(month))
  

stats_frame = plotting_master_frame %>%
  group_by(scenario,
           month,
           variable) %>%
  summarize(max    = max(value),
            min    = min(value),
            mean   = mean(value),
            p05    = quantile(x = value, probs=0.05),
            p25    = quantile(x = value, probs=0.25),
            p50    = median(value),
            p75    = quantile(x = value, probs=0.75),
            p95    = quantile(x = value, probs=0.95)) %>%
  ungroup() %>%
  mutate(month_num = as.numeric(month))
  
```



# Plot frame

```{r}


for (Targert_Variable in Variable) {
  
  plotting_frame = plotting_master_frame %>%
    filter(variable == Targert_Variable)

  ribbon_frame = stats_frame %>%
    filter(variable == Targert_Variable) 

  myplot = ggplot(data    = plotting_frame) +
            aes(x    = month_num,
                y    = value) +
            labs(title = "Point-based Thornthwaite Budget",
                 subtitle = str_c(baseline_period,  
                                  " Historical Period vs ",
                                  examined_period,
                                  " Future Period",
                                  sep = ""),
                 caption = str_c("Coordinates: ",
                                 target_longitude," °E, ",
                                 target_latitude, " °N",
                                 sep = ""), 
                 y     = Targert_Variable) +
            scale_fill_manual( values = c( "cyan",      
                                           "yellow",    
                                           "magenta"),
                               name   = NULL) + 
            scale_color_manual(values = c( "darkcyan", 
                                           "goldenrod", 
                                           "darkmagenta"),
                               name   = NULL) +
            theme_bw() +
            theme(strip.background = element_rect(fill=NA),
                  axis.title.x     = element_blank(),
                  axis.ticks.x     = element_blank()) +
            scale_x_continuous(breaks       = unique(plotting_frame$month_num),
                               labels       = unique(plotting_frame$month),
                               minor_breaks = NULL) +
            geom_ribbon(data    = ribbon_frame,
                        mapping = aes(x     = month_num,
                                      y     = NULL,
                                      ymax  = p95,
                                      ymin  = p05,
                                      color = NULL,
                                      fill  = scenario),
                          alpha = 0.1) +
            geom_ribbon(data    = ribbon_frame,
                        mapping = aes(x     = month_num,
                                      y     = NULL,
                                      ymax  = p75,
                                      ymin  = p25,
                                      color = NULL,
                                      fill  = scenario),
                          alpha = 0.2) +
          
            geom_line(data    = ribbon_frame,
                      mapping = aes(x     = month_num,
                                    y     = mean,
                                    color = scenario))  
    
    
  print(myplot)  
    
 }   
  
```
